---
title: "Lab 1"
author: "MGW"
date: "1/9/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(readr)
```

## Chunk 1

- Control Alt I opens a new R chunk.

- Here tells you to look in the following folder to find the file

```{r}
us_landings<-read.csv(here("data","noaa_fisheries.csv"))
```

## Chunk 2

- Now, make a tidy version of the data frame

- Janitor is cleaning up the dataframe

  - clean_names made everything lowercase
  - mutate: add an additional column that means something new
    - ..unless you want to change the existing variable!
    - put the existing name of the variable after mutate (in this case, state)
    - apply it to the existing variable, state (after str_to_lower)
    - this puts everything in lowercase!

```{r}
landings_tidy <- us_landings%>%
  janitor::clean_names()%>%
  mutate(state=str_to_lower(state),
         afs_name=str_to_lower(afs_name))
  
```

## Chunk 3

- In the readr package, there's a bunch of options for parsing things, such as parse number. If you only want to get the numeric version of an amount of money (such as only wanting 50 from $50), use parse number in readr.

```{r}
landings_tidy<-landings_tidy%>%
  mutate(dollars_num=parse_number(as.character(dollars_usd)))
```

## Chunk 4

- Remove the word "aggregate" from all rows in the column afs_name
- `str_detect(salmon_landings$afs_clean,pattern="salmon")` tells you if each row has the word 'salmon' in it or not for that variable
- Now, let's filter out every row with the word salmon into two groups using the command `separate`
- This created a new dataset called salmon_landings with only 1500 obs

```{r}
salmon_landings <- landings_tidy%>%
  mutate(afs_clean=str_remove(afs_name, pattern="aggregate"))%>%
  filter(str_detect(afs_clean, pattern="salmon"))%>%
  separate(afs_clean, into=c("group","species"),sep=",")
```

## Chunk 5

- Find some grouped summary data:

- Find annual total US landings and dollar value (summing across all states) for each type of salmon using `group_by` and `summary`

- Reduces to <500 rows because there's like 60 years and a certain number of species per year

```{r}
salmon_summary<-salmon_landings%>%
  group_by(year, species) %>%
  summarize(
    tot_landings = sum(landings_pounds),
    tot_value = sum(dollars_num)
  )
```

## Chunk 6

- Now, let's plot
- group gives different lines for each group (could also do color = species) in either the first aesthetic or the second aesthetic

```{r}
salmon_landings_graph <- ggplot(data=salmon_summary,
                                aes(x=year,y=tot_landings))+
  geom_line(aes(color=species),size=1)+
  theme_minimal()

salmon_landings_graph
```

## Chunk 7

- Export graph with certain dimensions for journal

```{r}
ggsave(plot=salmon_landings_graph,
       here("figures","mygraph.png"),
       height=5,width=8)
```

## Chunk 8

- Make a nice kable table:

```{r}
salmon_first_5 <- salmon_summary %>%
  head(5)

kable(salmon_first_5)%>%
  kable_styling(bootstrap_options="striped",
                full_width=FALSE)
```

